I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-17 Paper/proofs.tex
update proof

	\begin{definition}[Star and Sharp]
	\begin{enumerate}
		\item  $\vx^{*}$ is a input vector that for every node $a_i$, if $S(a_i)=1$, then $x_{a_i}$ is the maximal value of $a_i$, and if $S(a_i)=-1$, then $x_{a_i}$ is the minimal value of $a_i$ (if $S(a_i)=0$, then $x_{a_i}$ be an arbitrary fixed value in the input interval of $a_i$).

		\item  $\vx^{\sharp}$ is a input vector that for every node $a_i$, if $S(a_i)=-1$, then $x_{a_i}$ is the maximal value of $a_i$, and if $S(a_i)=1$, then $x_{a_i}$ is the minimal value of $a_i$ (if $S(a_i)=0$, then $x_{a_i}$ be an arbitrary fixed value in the input interval of $a_i$).
	\end{enumerate}

@@

				In this section, we will try to prove theorem 2, that is

				\begin{theorem}
					\label{no_diamond_3}

					Suppose we have a DNN with $\ReLU$ as the unique activation functions and with a unique output node $z$. We consider compensating pairs in between any two nodes (i.d., the source node or target node can in hidden layers.)

					Using MILP method, if every nodes $b$ in any compensating pair
					$(\pi,\pi')$ (except the start node and end node) is encoded as a binary/integer variable (opened), then the upper and lower 
					bounds computed by MILP are the exact max and min value of $z$.
				\end{theorem}

				Before the proof of above theorem, we will try to weaker one, that is, consider not only compensating pairs, but also pairs that allows common nodes (the two paths in a compensating pairs cannot have a common nodes). We call such pair a weak compensating pair.

				\begin{definition}
					We define a decomposition of the input nodes set to be $A_{pure}\sqcup A_{open}$ as follows:  
					\begin{itemize}
						\item $a_k \in A_{pure}$  if
						\begin{enumerate}
							\item every path from $a_k$ to $z$ has weight $\geq 0$, or
							\item every path from $a_k$ to $z$ has weight $\leq 0$.

@@

				\begin{lemma} \label{lem:open_node_2}
					A node $b$ in the hidden layer will not be on a weak compensating pair if and only if one of the following two happens:
					\begin{enumerate}
						\item Any path from $b$ to the output node $z$ has a weight $0$, and we use $w_{b,z}=0$ to denote this case, or
						\item For every input node $a_j\in J$, for any path from $a_j$ to $b$ has a weight of $0$. We use $w_{a_j,b}=0$ to denote this case.
					\end{enumerate}
					We denote $B_{zero}$ the set of such nodes $b$ that case 1 holds, and denote $B_{pure}$ the of such nodes $b$ that case 2 holds but case 1 does not hold.

					We denote $B_{open}$ the set of all nodes $b$ in a compensating path.
				\end{lemma}

@@

				\begin{lemma}\label{lem:subnetwork}
					$B_{pure}$ and $I$ can form a sub-network: every node in $B_{pure}\cup I$ only replies on nodes in $B_{pure}\cup I$ in previous layers. We call this sub-network $N_I$. In $N_I$, there is no compensating pair.
				\end{lemma}

@@

				Consider the sign of nodes function from Definition \ref{sign_of_nodes}. Because now we allow compensating paths, we cannot define this function over all nodes.

				\begin{definition}\label{sign_of_nodes_in_I_2}
					We define a partial sign function $S$ over nodes $n$ such that : 	
					\begin{enumerate} 
						\item all paths from $n$ to $z$ have 0 weight, and then $S(n)=0$; 
						\item all paths from $n$ to $z$ have non-negative weight, and at least one path has a positive weight, and then $S(n)=1$; 
						\item all path from $n$ to $z$ has non-positive weight, and at least one path has a negative weight, and then $S(n)=-1$.
						\item Otherwise $S(n)$ is undefined.
					\end{enumerate}
				\end{definition}

				Notice that $S$ is defined on all nodes $a_i \in I$ (it can also be defined trivially for all nodes $b$ as there is a unique path to $z$, and at $z$ also). And similarly, for nodes in  $B_{zero}$, $S$ is defined. We introduce the following lemma:

@@

				\begin{proof}
					Otherwise, suppose there exists two paths $P_1,P_2$ from $b$ to $z$ with (strictly) different signs. Then consider $\langle a_i,b\rangle+P_1$ and $\langle a_i,b\rangle+P_2$, they are two paths with weights with different signs. This is a contradiction since $a_i\in I$.
				\end{proof}
				Therefore, for any node $b$  that $S$ is undefined on $b$, it only depends on the values on $J$.

@@

				\begin{lemma}\label{lem:subnetwork2}
					The input set $J$ with nodes that only depending on $J$ forms a sub-network. This network contains all nodes that $S$ is not defined on them. We call this sub-network $N_J$.
				\end{lemma}

@@

				By definition, for any $b\in N_J$, either $b\in B_{open}$, or $b$ is a constant node.

				We denote $\boldsymbol{x}_S$ for any subset $S\subseteq K$ to refer the input vector $\langle x_{a_k}\rangle_{a_k\in S}$. We also denote $\boldsymbol{x}_I\oplus \boldsymbol{x}_J = \boldsymbol{x}_K$ and $\Val(z)=\Val_{\boldsymbol{x}_K}(z)=\Val_{\boldsymbol{x}_I,\boldsymbol{x}_J} (z)$. We can see that for a node $b$ that $S$ is not defined on it, if $\boldsymbol{x}_J$ is fixed, then the value of $b$ becomes a constant.

@@

					Assume $b$ is the output node or a node in a hidden layer that $S(b)=1$. Then
					$\max_{\boldsymbol{x}} (\Val_{\boldsymbol{x}}(b)) =\max_{\{\boldsymbol{x} \mid \boldsymbol{x}_I=\boldsymbol{x}^*_I\}} (\Val_{\boldsymbol{x}}(b))$.

>>>>

	\begin{definition}[Star and Sharp]

		\begin{itemize}
			\item $\vx^{*}$ is a input vector that for every node $a_i$, if $S(a_i)=1$, then $x_{a_i}$ is the maximal value of $a_i$, and if $S(a_i)=-1$, then $x_{a_i}$ is the minimal value of $a_i$ (if $S(a_i)=0$, then $x_{a_i}$ be an arbitrary fixed value in the input interval of $a_i$).

			\item  $\vx^{\sharp}$ is a input vector that for every node $a_i$, if $S(a_i)=-1$, then $x_{a_i}$ is the maximal value of $a_i$, and if $S(a_i)=1$, then $x_{a_i}$ is the minimal value of $a_i$ (if $S(a_i)=0$, then $x_{a_i}$ be an arbitrary fixed value in the input interval of $a_i$).
		\end{itemize}

@@

				In this section, we will prove theorem \ref{th2}. We will prove that theorem by proving the following theorem:

				\begin{theorem}

					Suppose $\mathcal{N}$ is a DNN with $\ReLU$ as the unique activation functions and with a unique output node $z$.

					Using MILP method, if every nodes $b$ in any compensating pair
					$(\pi,\pi')$ (except the start node and end node) is encoded as a binary/integer variable (opened), then the upper and lower 
					bounds computed by MILP are the exact maximal and minimal value of $z$.
				\end{theorem}

				\subsection{Preparation}

				Before the proof of above theorem, we will prove a weaker version, that is, consider not only compensating pairs, but also pairs that allows common nodes (the two paths in a compensating pairs cannot have a common nodes). We call such pair a weak compensating pair.

				\begin{definition}
					We define a decomposition of the input nodes set $A_{pure}\sqcup A_{open}$ as follows:  
					\begin{itemize}
						\item $a_k \in A_{pure}$  if either
						\begin{enumerate}
							\item every path from $a_k$ to $z$ has weight $\geq 0$, or
							\item every path from $a_k$ to $z$ has weight $\leq 0$.

@@

				\begin{lemma} \label{lem:open_node_2}
					A node $b$ in the hidden layer will not be on a weak compensating pair if and only if one of the following two happens:
					\begin{enumerate}
						\item Any path from $b$ to the output node $z$ has a weight $0$, and we use $w_{b,z}=0$ to denote this case, or
						\item For every input node $a_j\in J$, every path from $a_j$ to $b$ has a weight of $0$. We use $w_{a_j,b}=0$ to denote this case.
					\end{enumerate}

					We denote $B_{zero}$ the set of such nodes $b$ that case 1 holds, and denote $B_{pure}$ the of such nodes $b$ that case 2 holds but case 1 does not hold.	We denote $B_{open}$ the set of all nodes $b$ in a compensating path.

@@

				\begin{lemma}\label{lem:subnetwork}
					$B_{pure}$ and $I$ can form a sub-network: every node in $B_{pure}\cup I$ only depends on nodes in $B_{pure}\cup I$ backward: if $c$ is in $B_{pure}\cup I$ but $b$ is not, then $W_{bc}=0$. We call this sub-network $N_I$. In $N_I$, there is no compensating pair.
				\end{lemma}

@@

				Recall the sign of nodes function from Definition \ref{sign_of_nodes}. Because now we allow compensating paths, we cannot define this function over all nodes.

				\begin{definition}\label{sign_of_nodes_in_I_2}
					We define a partial sign function $S$ over nodes $n$ such that : 	
					\begin{enumerate} 
						\item if all paths from $n$ to $z$ have 0 weight, then $S(n)=0$; 
						\item if all paths from $n$ to $z$ have non-negative weight, and at least one path has a positive weight, then $S(n)=1$; 
						\item if all path from $n$ to $z$ has non-positive weight, and at least one path has a negative weight, then $S(n)=-1$.
						\item Otherwise $S(n)$ is undefined.
					\end{enumerate}
				\end{definition}

				Notice that $S$ is defined on all nodes $a_i \in I$ (it can also be defined trivially for all nodes $b$ as there is a unique path to $z$, and at $z$ also). For the domain of $S$, we introduce the following lemma:

@@

				\begin{proof}
					Otherwise, suppose there exists two paths $P_1,P_2$ from $b$ to $z$ with (strictly) different signs. Then consider $\langle a_i,b\rangle+P_1$ and $\langle a_i,b\rangle+P_2$, they are two paths with weights with different signs. This is a contradiction since $a_i\in I$.
				\end{proof}
				Therefore, for any node $b$  that $S$ is undefined on $b$, it only depends on input nodes in $J$.

@@

				\begin{lemma}\label{lem:subnetwork2}
					The input set $J$ with nodes that only depends on $J$ forms a sub-network. This network contains all nodes that $S$ is not defined on them. We call this sub-network $N_J$.
				\end{lemma}

@@

				By definition, for any $b\in N_J$, either $b\in B_{open}$, or $b$ is a constant node: to it, any path from any input node must have zero weight.

				We denote $\boldsymbol{x}_S$ for any subset $S$ of input nodes to refer the input vector $\langle x_{a_k}\rangle_{a_k\in S}$. We also denote $\boldsymbol{x}_I\oplus \boldsymbol{x}_J = \boldsymbol{x}$ and $\Val(z)=\Val_{\boldsymbol{x}}(z)=\Val_{\boldsymbol{x}_I,\boldsymbol{x}_J} (z)$. For a node $b$ that $S$ is not defined on it, since $b$ only depends on $J$, if $\boldsymbol{x}_J$ is fixed, then the value of $b$ will be fixed.

@@

					Assume $b$ is the output node or a node in hidden layers that $S(b)=1$, then
					$\max_{\boldsymbol{x}} (\Val_{\boldsymbol{x}}(b)) =\max_{\{\boldsymbol{x} \mid \boldsymbol{x}_I=\boldsymbol{x}^*_I\}} (\Val_{\boldsymbol{x}}(b))$.
